---
output: 
  pdf_document:
    citation_package: natbib
    keep_tex: true
    fig_caption: true
    latex_engine: pdflatex
    template: template.tex
    number_sections: false
title: "Estimated parameters for Schistosomiasis in Sub-saharan Africa"
# thanks: "**Corresponding author**: c.fronterr@lancaster.ac.uk."
# author:
#   name: Claudio Fronterre, Luigi Sedda
#   affiliation: CHICAS, Lancaster University
# abstract: "This is my abstract"
# keywords: "keyword1, keyword2, Keyword3"
date: "`r format(Sys.time(), '%B %d, %Y')`"
geometry: margin=1in
fontfamily: mathpazo
fontsize: 11pt
header-includes:
- \usepackage[english]{babel}
- \usepackage{booktabs}
- \usepackage{longtable}
- \usepackage{array}
- \usepackage{multirow}
- \usepackage[table]{xcolor}
- \usepackage{wrapfig}
- \usepackage{float}
- \usepackage{colortbl}
- \usepackage{pdflscape}
- \usepackage{tabu}
- \usepackage{threeparttable}
- \usepackage{threeparttablex}
- \usepackage[normalem]{ulem}
- \usepackage{makecell}
#spacing: double
bibliography: biblio.bib
biblio-style: apalike
endnote: no
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F, cache = F, warning = F, message = F)
library(ggplot2)
library(dplyr)
source("R/functions.R")
```

# Data processing

ESPEN data was checked for the presence of duplicated locations. Only locations 
replicated over time where found and, for each of these entries, the most recent
survey was retained. All the countries with less than 50 surveys were excluded
from modelling.

# Modelling framework

We fit a binomial geostatistical model to Schistosomiasis data for each country 
both for Haematobium and for Mansoni. The number of individuals, $y_i$, tested positive for Schistosomiasis out of the number of people examined $n_i$ at each location $x_i$ can be considered as the realisation of a Binomial random variable $Y_i \sim \text{Binom}(n_i, p(x_i))$, with  $p(x_i)$ modelled as: 
$$\log\left\{\frac{p(x_i)}{1 - p(x_i)}\right\}=\mu+S(x_i) + Z_i$$
where $S(x)$ is a zero mean Gaussian process with variance $\sigma^2$ and an exponential correlation function given by $\rho(u;\phi)=\exp\{-u/\phi\}$ where $\phi > 0$ is a scale parameter that controls the extent of the spatial correlation and $u$ is the distance between two sampling locations. Moreover, $Z_i$ is a set of independent zero-mean Gaussian variables with mean $0$ and variance $\tau^2$. 

# Results

\autoref{tab:haema} and \autoref{tab:mansoni} report the estimated parameters obtained by fitting a binomial geostatistical model to Haematobium and Mansoni pre-intervention data
respectively. In \autoref{fig:plots} we compare the distribution of the estimated model parameteres for the two species.

```{r tab_haema}
tab_haema <- readr::read_csv("output/models_fitted/haema/fitted_parameters.csv") 

knitr::kable(tab_haema, digits = 3, row.names = F,
             col.names = c("Country", "n", "$\\hat{\\mu}$", "$\\hat{\\sigma}^2$",
                           "$\\hat{\\phi}$", "$\\hat{\\tau}^2$"), 
             escape = F, align = "lcccc",
             caption = "\\label{tab:haema} Monte Carlo maximum likelihood 
                        estimates for the model fitted to Haematobium data.",
             )
```

```{r tab_masoni}
tab_mansoni <- readr::read_csv("output/models_fitted/mansoni/fitted_parameters.csv") 

knitr::kable(tab_mansoni, digits = 3, row.names = F,
             col.names = c("Country", "n", "$\\hat{\\mu}$", "$\\hat{\\sigma}^2$",
                           "$\\hat{\\phi}$", "$\\hat{\\tau}^2$"), 
             escape = F, align = "lcccc",
             caption = "\\label{tab:mansoni} Monte Carlo maximum likelihood 
                        estimates for the model fitted to Mansoni data.",
             )
```

```{r plots, fig.align='center', out.width="100%", fig.cap="Distribution of parameters estimates for both species."}
tab_haema$specie <- "Haematobium"
tab_mansoni$specie <- "Mansoni"

parameters <- rbind(tab_haema, tab_mansoni)

parameters %>% 
  select(-country, -n) %>% 
  tidyr::gather(key = "parameter", value = "estimate", -specie) %>% 
  ggplot(aes(x = estimate, fill = specie)) +
  geom_histogram(position = position_identity(), alpha = .7, bins = 24) +
  facet_wrap(~ parameter, scales = "free", labeller = label_parsed) +
  scale_fill_brewer(type = "qual", palette = 6) +
  labs(x = "Estimated value", fill = "Specie", y = "") + 
  theme_bw() +
  theme(legend.position = "top")
```

